### 2019-10-14 paper/SortingAlgorithms.tex

\section{Sorting Algorithms}

This work uses four sorting algorithms: quicksort, mergesort, insertionsort, and bubblesort. In the following subsections, we'll give an overview of them.

\subsection{Quicksort}

Quicksort is an algorithm whose execution time is $\theta(n^2)$ in the worst case over \textit{n} numbers as input. Despite that execution time, quicksort is often the best option for sorting because of its remarkable average efficiency: $\theta(n lg n)$ \cite{Cormen2009}.

\subsection{Mergesort}

\subsection{Insertionsort}

\subsection{Bubblesort}