### 2019-10-21 paper/Abstract.tex
continuação da escrita do background

\begin{abstract}
Although most people in the world use technology devices for many tasks, they don't know how the devices work and how they deal with faults. When those faults occur in memory, software behavior could be affected. Together with the software-specific algorithms are the sorting algorithms used to solve problems like ordering a list of products by their price. This work presents a discussion about how quicksort, mergesort, insertionsort and bubblesort algorithms are affected by memory faults.
\end{abstract}

\begin{IEEEkeywords}
sorting algorithms, memory faults.
\end{IEEEkeywords}