### 2019-12-20 paper/Background.tex

\section{Background}

In this section we describe basic concepts about memory faults and sorting algorithms.

\input{MemoryFaults}

\input{SortingAlgorithms}